---
title: 'Exploring Predictive Factors for Heart Disease'
author: "Jiali Guo"
date: "`r Sys.Date()`"
output: html_document
params: 
  production: TRUE
---

Refers to 
https://www.kaggle.com/datasets/yasserh/heart-disease-dataset/data


```{r, set-chunk-opts, echo = FALSE, eval = params$production}
# set global chunk options
# eval -- run the code or not
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)
```

## Introduction
This dataset comprises records of individuals monitored for heart health, presenting a binary classification problem aimed at predicting the presence or absence of heart disease. Each record includes various demographic and clinical variables known to be associated with cardiovascular health.
This dataset contains `r nrow(data)` rows and `r ncol(data)` columns.

__Objective__
<ol>
  <li> Explore the Demographic Profile: Investigate the distribution of age and gender within the sample population to understand the demographic characteristics of the heart disease patients.</li>
  <li> Evaluate Risk Factors: Assess relationship between chest pain type, resting blood pressure, serum cholesterol level, and fasting blood sugar levels with the presence of heart disease. Determine if these factors serve as potential risk indicators.</li>
</ol>


```{r, load-data}
here::i_am("report.Rmd") #identify the project root directory
data <- readRDS(
  file = here::here("output/data_clean.rds")
)
```

## Descriptive Analysis

__Table 1. Descriptive statistics stratified by heart disease status__
```{r, table1}
table_one <- readRDS(
  file = here::here("output/table_one.rds")
)
table_one
```

The average age for heart disease patients is 52.5-year-old, which is slightly younger than healthy people (56.6-year-old) in this sample. The percentage of male subjects (68.3%) is twice the percentage of female (31.7%) subjects. Female heart disease patients (72) are much more than healthy females (24). 42% of patients have type 3 chest pain, while only 13% of healthy people have the same problem. There is no significant different in resting blood pressure and serum cholesterol level between healthy and heart disease subjects. However, 76 healthy people report exercise induced angina and only 23 heart disease patients have this problem.


## Fitting logistic regression model

I fit generalized linear model for this binary outcome. When a binary outcome variable is modeled using logistic regression, it is assumed that the logit transformation of the outcome variable has a linear relationship with the predictor variables.
```{r, logistic-regression-model}
logistic_regression_table <- readRDS(
  file = here::here("output/logistic_regression_table.rds")
)
logistic_regression_table
```

```{r, odds-ratio-table}
odds_ratio_table <- readRDS(
  file = here::here("output/odds_ratio_table.rds")
)
odds_ratio_table

```
The odds of having heart disease increase by a factor of 1.77 (95% CI = [0.58, 5.77]) for people with fasting blood sugar > 120 mg/dl versus those fasting blood sugar <= 120 mg/dl.

People experiencing exercise-induced angina have a 53% lower odds of having heart disease compared to those without, with a 95% confidence interval ranging from 0.20 to 1.14.

Females have a 81% lower odds of having heart disease compared to males, with a 95% confidence interval ranging from 0.06 to 0.54.


## Figures

### Bar Chart
```{r, bar-chart}
knitr::include_graphics(
  here::here("output/barchart.png")
)
```

Males are more likely to have heart disease than females. In this sample, around 100 males have heart disease and merely 30 females have heart disease.


### Violin Plot
```{r, violin-plot}
knitr::include_graphics(
  here::here("output/violinplot.png")
)
```

Despite different chest pain types, the serum cholesterol level mainly concentrates in 230 mg/dl in this sample population. For people who have type 2 chest pain, their serum cholesterol level are more widely distributed, ranging from 120 mg/dl to 550 mg/dl.

### Histogram
```{r, histogram}
knitr::include_graphics(
  here::here("output/histogram.png")
)
```

Resting blood pressure concentrates around 120 mm Hg ~ 140 mm Hg.
